Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure

### 2017-04-06 documents/TriangleScenarioArXiv/ts.tex
compatibility section no-longer references marginal models

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    \affiliation{University of Waterloo, Waterloo, Ontario, Canada, N2L 3G1}
    \author{Elie Wolfe}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    % \email{ewolfe@perimeterinstitute.ca}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are often incompatible with particular assumption about classical causal structure. Such correlations are sometimes termed non-local, although they are more accurately termed non-classical, as the presence of correlations certifies the genuinely non-classical nature of a causal setup in a device-independent fashion.  Whenever a quantum correlation violates a causal compatibility inequality (e.g. a Bell inequality) then the correlation is witnessed as non-classical. Any constraint which is satisfied by all correlations which \textit{are} classically compatible with a given causal structure defines a causal compatibility criterion. Such criteria were recently were recently derived for the Triangle structure [\href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}] in the form of polynomial inequalities, begging the question: do any of those inequalities admit violation by quantum correlations? Numerical investigation suggests not, and we further conjecture that the set of correlations admitted by the classical Triangle structure is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution in this work, however, is the derivation of new causal compatibility inequalities for the Triangle structure which \textit{do} admit quantum violation. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    \tableofcontents
    \setlength{\parskip}{1em}
    \clearpage

    \section{Introduction}
    \label{sec:introduction}
    In recent decades, the technical utility of quantum mechanics has become abundantly clear. In the realm of computation, quantum algorithms -- such as Shor's algorithm~\cite{Shor_1997} and numerous others~\cite{Jordan_2016} -- scale exponentially better than their classical counterparts. In the realm of secure communication, quantum protocols --- a popular example being Quantum Key Distribution~\cite{Bennett_2014} -- are able to provide privacy even against hypothetical adversaries with unlimited computational power, a desideratum which classical protocols are unable to fulfill. History, every opportunity of quantum phenomena which fails to be emulated by classical resources has led to a practical exploitation of the non-classical feature in order to solve a some computational of communicational problem~\cite{Neilsen_Chaung_2011}. Motivated by past successes, a primal objective of modern quantum information theory is the discovery of new situations wherein quantum mechanics offers an advantage, and certification that the quantum advantage is genuine.

    From a foundational prospective, the most robust demonstrations of quantum phenomena with no classical emulation have involved Bell inequalities~\cite{Bell_1964}. Originally, Bell inequalities were derived as a way to show that no hidden variable theory could ever account for quantum mechanics; in this sense Bell inequalities are a response to the famous EPR paradox~\cite{EPR_Orig}. The enumeration of Bell inequalities has since become a widespread systematic way of demonstration the \textit{non-classicality} of a given observation. More recently it has been appreciated that Bell inequalities can understood as consequences of \term{causal inference}~\cite{Wood_2012}. Causal inference is concerned with classifying observations into those which can and cannot be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Causal compatibility inequalities, such as Bell of Instrumental inequalities, characterizer the space of observations that are compatible with a hypothesized causal structure, albeit the characterization offered by practically derivable inequalities often only an approximation. To derive the traditional Bell inequalities from causal inference one starts with a (classical) causal structure known as the Bell structure, as depicted in \cref{fig:bell_scenario}. The fundamental Bell structure involves non-communicating parties making measurements on some hidden shared resource $\la$, where the measurement outcomes ($A, B$) are presumed to be stochastic functions of the the local choices of measurement settings ($S_A, S_B$) and the shared resource $\la$. Quantum non-classicality in the Bell structure has been thoroughly studied since Bell's original work~\cite{Brunner_2013}. More complex scenarios, however, such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014}, are much less understood. Here we investigate one particular causal structure: the Triangle structure (\cref{fig:triangle_structure}).

    The Triangle structure (\cref{fig:triangle_structure}) is a causal structure composed of $3$ parties labeled $A, B, C$ arranged in a triangular configuration while pair-wise sharing hidden/latent variables $X, Y, Z$. It has been extensively studied previously (see~\cite[Fig. 1]{Steudel_2010},~\cite[Fig. 6]{Chaves_2014},~\cite[Fig. 8]{Branciard_2012},~\cite[Fig. 8, App. E]{Henson_2014},~\cite[Fig. 3]{Fritz_2012},~\cite[Fig. 1]{Inflation}, $\ldots$). An overview of some milestone results is provided in \cref{sec:triangle_structure}. Identifying causal compatibility inequalities for this configuration has been seen as particularly challenging~\citet{Branciard_2012}. Further identifying causal compatibility inequalities of such high resolution that such inequalities can be violated by quantum-accessible distributions has remained out-of-reach for the Triangle structure.

@@

    The manuscript is organized into the following sections: \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:triangle_structure} discusses the Triangle structure and provides an overview of existing research; identifying its stark differences from the Bell structure and motivating why the Triangle structure was chosen as a case study. \Cref{sec:fritz_distribution} defines and discusses a singularly-quantum correlation first conceived of by \citet{Fritz_2012}, which we term the \term{Fritz distribution}. The Fritz distribution was proven to be non-classical without the use of inequalities --- this report fills in the desired missing inequalities. \Cref{sec:inflation_technique} briefly summaries the Inflation Technique in the specific context of this work. Although the summary presented here is designed to be self-standing, a much more pedagogical introduction is offered by the original work~\cite{Inflation}. \Cref{sec:deriving_inequalities} demonstrates how the inflation technique can be used to derive causal compatibility inequalities for the Triangle structure that are violated by quantum-accessible distributions. A sample of these inequalities are presented at the end of \cref{sec:found_inequalities} (specifically \cref{eq:ww_ineq}) as well as in \cref{sec:web_inequality,sec:symmetic_web_inequality}. Finally, \cref{sec:violations_noise} introduces the method for parameterizing the space of quantum-accessible distributions which we used to numerically optimize against our derived inequalities in search of stronger types of non-classicality. Additionally, the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities.

    \section{Causal Compatibility}
    \todo[TC]{Talk to Elie about the varying notions of compatibility}
    \label{sec:causal_compatibility}
    The task of causal inference aims to determine the set of potentially observable probability distribution compatible with some hypothesis about causal relationships. whether or not a given family of observed probability distributions can be explained by a proposed causal mechanism~\cite{Pearl_2009}. If a set of observed distributions can be explained by the hypothesized causal mechanism, then the observations are said to be \textit{compatible} with the causal mechanism. In order to define compatibility rigorously, we first need to formally define what is meant by ``set of observations''.

    \subsection{Marginal Models}
    Given a set of observable variables $\jointvar = \bc{v_1, \ldots, v_k}$ where each $v \in \jointvar$ is a statistical random variable, one can conceivably make a measurement over of these variables and obtain a joint probability distribution $\prob[\jointvar] = \prob[v_1,\ldots,v_n]$. More generally however, and depending on the context, it is possible to make a measurement on a set of the system $V \subset \jointvar$ in order to obtain a marginal distribution $\prob[V]$. With complete generality however, one could in principle be limited so as to only be able to record observations on portions of the system $\mscenario = \bc{V_1, \ldots, V_m \mid \forall i : V_i \subseteq \jointvar}$ obtaining a collection of marginal probability distributions $\bc{\prob[V_1], \ldots, \prob[V_m] \mid \forall i: V_i \in \mscenario}$.

    \subsection{Causal Structure}
    A hypothesis of causal mechanism is formally referred to as  a \term{causal structure}. Mathematically, a causal structure can be represented as a directed acyclic graph $\graph = \br{\nodes, \edges}$ of nodes $\nodes$ and edges $\edges$ as defined in \cref{sec:graph_theory}. The nodes of a causal structure are random variables and are classified into two disjoint classes: the \term{latent nodes} $\nodes_L$ representing the set of nodes that are unobservable and the \term{observed nodes} $\nodes_O$. The latent nodes of a causal structure represent completely hidden variables that are hidden by some fundamental process or cannot be measured due to other limitations. The observed nodes represent random variables that are measurable and in practice will be observed by the marginal model $\prob^\mscenario$. The edges $\edges$ of a causal structure represent casual influence between variables within the causal structure.

    A number of causal structures are depicted throughout this document (see \cref{fig:bell_scenario,fig:triangle_structure} as well as \cref{fig:inflations}). As an example, the Triangle structure has $3$ observable nodes and $3$ latent nodes,
    \[ \nodes_{O} = \bc{A, B, C} \qquad \nodes_{L} = \bc{X, Y, Z} \]
    along with $6$ edges indicating that the latent nodes $\bc{X, Y, Z}$ cause and influence the observed nodes $\bc{A, B, C}$.
    \[ \edges = \bc{\bc{X \to A}, \bc{Y \to A}, \bc{Y \to B}, \bc{Z \to B}, \bc{Z \to C}, \bc{X \to C}} \]

    \subsection{Compatibility}
    \label{sec:compatibility}

    Now that the notions of a causal structure $\graph$ and a marginal model $\prob^{\mscenario}$ are defined, it is now possible to precisely define the notion of compatibility between the two. Intuitively, the absence of certain causal connections within $\graph$ should restrict and limit the set of possible observations that can be made on $\graph$; equivalently, the causal connections within $\graph$ should explain $\prob^{\mscenario}$. If each variable $n \in \nodes$ is caused by its parents $\Pa[\graph]{n}$, then $n$ should be independent of all other variables that are not its descendants. This is known as the \term{local Markov property}:
    \[ n \indep \nodes \setminus \De[\graph]{n} \mid \Pa[\graph]{n} \quad \forall n \in \nodes \eq \label{eq:local_markov_property} \]
    Where $\indep$ denotes statistical independence. With this as motivation, we define a set of \term{causal parameters} for a causal structure $\graph$ to be a specification of a conditional distribution for every variable $n \in \nodes$ given its parents:
    \[ \bc{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes} \eq \label{eq:causal_parameters} \]
    It is important to note that a set of causal parameters is not unique; the numerical nature of set of causal parameters is a \text{choice}. Altogether, a marginal model $\prob^{\mscenario}$ is \term{compatible} with a causal structure $\graph$ if there exists a choice of causal parameters such that each context $\prob[V] \in \prob^{\mscenario}$ can be \textit{recovered} from the following series of operations:
    \begin{enumerate}
        \item Obtain a joint probability distribution over \textit{all} variables of the causal structure using the chosen casual parameters:\footnote{Here $\prod$ denotes the product distribution $\prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] = \prob[n_1 \mid \Pa[\graph]{n_1}] \times \cdots \times \prob[n_k \mid \Pa[\graph]{n_k}]$.}
        \[ \prob[\nodes] = \prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] \eq \label{eq:local_markov_property_operation}\]
        This step ensures that the local Markov property (\cref{eq:causal_parameters}) holds for all variables $n \in \nodes$.
        \item Marginalize over the latent variables of $\graph$ to obtain a joint distribution over the observed variables $\nodes_O$:
        \[ \prob[\nodes_O] = \sum_{\nodes_L} \prob[\nodes] \eq \label{eq:observable_marginalization} \]
        \item Finally marginalize over the observed nodes not in $V$ to obtain $\prob[V]$:
        \[ \prob[V] = \sum_{\nodes_O \setminus V} \prob[\nodes_O] \eq \label{eq:context_marginalization} \]
    \end{enumerate}
    It is important to note that compatibility between $\prob^{\mscenario}$ and $\graph$ requires that it is possible to recover \textit{each} context $\prob[V] \in \prob^{\mscenario}$ from the \textit{same} set of causal parameters. If it is not possible for $\prob^{\mscenario}$ to be recovered for \textit{any} choice of causal parameters, then $\prob^{\mscenario}$ is \term{incompatible}.

    \subsection{Inequalities}
    As defined in \cref{sec:compatibility}, determining compatibility is equivalent to determining if a set of causal parameters exists with the above-mentioned properties. In practice however, checking causal parameters systemically is intractable, especially when the cardinality (number of outcomes) of the latent variables is unspecified. Instead, \term{causal compatibility inequalities}\footnote{We refer to these inequalities as ``causal compatibility inequalities'' instead of simply Bell inequalities for two reasons. Firstly, ``Bell inequalities'' usually are associated specifically with the Bell structure/Bell causal structure. Secondly, the inequalities derived in this work are fundamentally distinct from a typical Bell inequality in that these inequalities are \textit{polynomial} over $\prob^{\mscenario}$ instead of \textit{linear}.} are probabilistic inequalities defined over the marginal contexts $\prob[V] \in \prob^{\mscenario}$ within the marginal model. If a marginal model $\prob^{\mscenario}$ happens to \textit{violate} any causal compatibility inequality, then that marginal model is definitely \textit{incompatible}. In \cref{sec:deriving_inequalities}, a brief discussion of existing techniques for deriving these inequalities will take place.

    Unfortunately, a singular inequality can not prove that a given marginal model is \textit{compatible}, only \textit{incompatible}. Instead, a \textit{complete characterization} of compatibility would consist of a complete set of all valid causal compatibility inequalities; satisfaction of a complete set of inequalities does prove compatibility. Currently however, it is unknown how to obtain a complete characterization for any given causal structure, including the Triangle structure.

    From the perspective of identifying quantum non-classicality, the causal structure $\graph$ acts as the classical hypothesis for a particular marginal model $\prob^{\mscenario}$. Therefore, non-classicality becomes synonymous with incompatibility: if a marginal model $\prob^{\mscenario}$ is incompatible, then it is non-classical. Henceforth, we will use these two terms interchangeably. From a resource standpoint, if the non-classicality of $\prob^{\mscenario}$ can be witnessed by an inequality $I$, but $\prob^{\mscenario}$ can be implemented using quantum states and measurements, then the causal compatibility inequality $I$ represents a concrete task where quantum resources out-perform classical resources.

@@

    \section{Symmetric Inequalities}
    \label{sec:symmetric_inequalities}

@@

    \subsection{Satisfaction of d-Separation Relations}
    \todo[TC]{Absorb the relevant components of the next section into this one}

    \subsection{Marginal Satisfiability and Inequalities}
    \label{sec:deriving_inequalities}

    In \cref{sec:causal_compatibility}, we defined the notion of a causal compatibility inequality without providing a means for deriving/computing them. This section aims to delineate algorithms and methods that can be used to derive causal compatibility inequalities for generic causal inference problems. Recalling the definition of causal compatibility presented in \cref{sec:compatibility}, a marginal model $\prob^{\mscenario}$ is compatible with a causal structure $\graph = \br{\nodes, \edges}$ if it can be recovered from a series of operations. First choose a set of causal parameters $\bc{\prob[n\mid\Pa[\graph]{n}] \mid n \in \nodes}$ and then construct a joint distribution $\prob[\nodes] = \prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}]$ over all nodes $\nodes$ (this step is \cref{eq:local_markov_property_operation}). Afterwards, each marginal distribution $\prob[V] \in \mscenario$ should be recoverable from a marginalization over $\prob[\nodes]$ (this step is the combination of \cref{eq:observable_marginalization,eq:context_marginalization}). Intuitively, these two requirements for a compatible marginal model $\prob^{\mscenario}$ provide two necessary components of compatibility:
    \begin{enumerate}
        \item The marginal model $\prob^{\mscenario}$ must admit a joint distribution $\prob[\nodes]$ (\cref{eq:observable_marginalization,eq:context_marginalization}).
        \item The joint distribution $\prob[\nodes]$ must satisfy the Markov separability criteria (\cref{eq:local_markov_property_operation}).
    \end{enumerate}
    In practice, this is the natural order in which to classify the compatibility of a marginal model $\prob^{\mscenario}$. First check if the marginal model admits a joint distribution and if and only if it does, determine whether or not the local Markov property can be satisfied for some choice of causal parameters. Intuitively, there are two reasons why a given marginal model $\prob^{\mscenario}$ would be deemed incompatible. One possibility is that the marginal model $\prob^{\mscenario}$ does not admit a joint distribution $\prob[\nodes]$. Alternatively, it could be possible that a joint distribution $\prob[\nodes]$ exists but can not factor into causal parameters as per \cref{eq:local_markov_property_operation}.

>>>>

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    \affiliation{University of Waterloo, Waterloo, Ontario, Canada, N2L 3G1}
    \author{Elie Wolfe}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    % \email{ewolfe@perimeterinstitute.ca}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are often incompatible with particular assumption about classical causal structure. Such correlations are sometimes termed non-local, although they are more accurately termed non-classical, as the presence of correlations certifies the genuinely non-classical nature of a causal setup in a device-independent fashion.  Whenever a quantum correlation violates a causal compatibility inequality (e.g. a Bell inequality) then the correlation is witnessed as non-classical. Any constraint which is satisfied by all correlations which \textit{are} classically compatible with a given causal structure defines a causal compatibility criterion. Such criteria were recently were recently derived for the Triangle structure [\href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}] in the form of polynomial inequalities, begging the question: do any of those inequalities admit violation by quantum correlations? Numerical investigation suggests not, and we further conjecture that the set of correlations admitted by the classical Triangle structure is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution in this work, however, is the derivation of new causal compatibility inequalities for the Triangle structure which \textit{do} admit quantum violation. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    \tableofcontents
    % \setlength{\parskip}{1em}
    \clearpage

    \section{Introduction}
    \label{sec:introduction}
    In recent decades, the technical utility of quantum mechanics has become abundantly clear. In the realm of computation, quantum algorithms --- such as Shor's algorithm~\cite{Shor_1997} and numerous others~\cite{Jordan_2016} --- scale exponentially better than their classical counterparts. In the realm of secure communication, quantum protocols --- a popular example being Quantum Key Distribution~\cite{Bennett_2014} --- are able to provide privacy even against hypothetical adversaries with unlimited computational power, a desideratum which classical protocols are unable to fulfill. Throughout history, every aspect of quantum phenomena which fails to be emulated by classical resources has led to a practical exploitation of the non-classical feature in order to solve a some computational of communicational problem~\cite{Neilsen_Chaung_2011}. Motivated by past successes, a primal objective of modern quantum information theory is the discovery of new situations wherein quantum mechanics offers an advantage, and certification that the quantum advantage is genuine.

    From a foundational prospective, the most robust demonstrations of quantum phenomena with no classical emulation have involved Bell inequalities~\cite{Bell_1964}. Originally, Bell inequalities were derived as a way to show that no hidden variable theory could ever account for quantum mechanics; in this sense Bell inequalities are a response to the famous EPR paradox~\cite{EPR_Orig}. The enumeration of Bell inequalities has since become a widespread systematic way of demonstration the \textit{non-classicality} of a given observation. More recently it has been appreciated that Bell inequalities can understood as consequences of \term{causal inference}~\cite{Wood_2012}. Causal inference is concerned with classifying observations into those which can and cannot be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Causal compatibility inequalities, such as Bell and Instrumental inequalities, characterize the space of observations that are compatible with a hypothesized causal structure, albeit the characterization offered by practically derivable inequalities is often only an approximation. To derive the traditional Bell inequalities from causal inference one starts with a (classical) causal structure known as the Bell structure, as depicted in \cref{fig:bell_scenario}. The fundamental Bell structure involves non-communicating parties making measurements on some hidden shared resource $\la$, where the measurement outcomes ($A, B$) are presumed to be stochastic functions of the the local choices of measurement settings ($S_A, S_B$) and the shared resource $\la$. Quantum non-classicality in the Bell structure has been thoroughly studied since Bell's original work~\cite{Brunner_2013}. More complex structures, however, such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014}, are much less understood. Here we investigate one particular causal structure: the Triangle structure (\cref{fig:triangle_structure}).

    The Triangle structure (\cref{fig:triangle_structure}) is a causal structure composed of $3$ parties labeled $A, B, C$ arranged in a triangular configuration while pair-wise sharing hidden/latent variables $X, Y, Z$. It has been extensively studied previously (see~\cite[Fig. 1]{Steudel_2010},~\cite[Fig. 6]{Chaves_2014},~\cite[Fig. 8]{Branciard_2012},~\cite[Fig. 8, App. E]{Henson_2014},~\cite[Fig. 3]{Fritz_2012},~\cite[Fig. 1]{Inflation}, $\ldots$). An overview of some milestone results is provided in \cref{sec:triangle_structure}. Identifying causal compatibility inequalities for this configuration has been seen as particularly challenging~\cite{Branciard_2012}. Further identifying causal compatibility inequalities of such high resolution that such inequalities can be violated by quantum-accessible distributions has remained out-of-reach for the Triangle structure.

@@

    The manuscript is organized into the following sections: \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:bell_structure} offers a summary of the popular Bell structure and associated inequalities. \Cref{sec:triangle_structure} discusses the Triangle structure and provides an overview of existing research; identifying its stark differences from the Bell structure and motivating why the Triangle structure was chosen as a case study. \Cref{sec:fritz_distribution} defines and discusses a singularly-quantum correlation first conceived of by \citet{Fritz_2012}, which we term the \term{Fritz distribution}. The Fritz distribution was proven to be non-classical without the use of inequalities --- this report fills in the desired missing inequalities. A sample of these inequalities are presented in \cref{sec:found_inequalities} (specifically \cref{eq:ww_ineq} as well as in \cref{sec:web_inequality,sec:symmetic_web_inequality}\todo[TC]{Reorganize the location of the inequalities}). \Cref{sec:inflation_technique} briefly summaries the Inflation Technique in the specific context of this work. Although the summary presented here is designed to be self-standing, a much more pedagogical introduction is offered by the original work~\cite{Inflation}. \Cref{sec:deriving_inequalities} demonstrates how the inflation technique can be used to derive causal compatibility inequalities for the Triangle structure that are violated by quantum-accessible distributions. Finally, \cref{sec:violations_noise} introduces the method for parameterizing the space of quantum-accessible distributions which we used to numerically optimize against our derived inequalities in search of stronger types of non-classicality. Additionally, the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities. \Cref{sec:conclusions} concludes with proposed avenues for further research.

    \section{Causal Compatibility}
    \label{sec:causal_compatibility}
    The task of causal inference is to determine the set of potentially observable probability distributions compatible with some hypothesis about causal relationships~\cite{Pearl_2009}. If an observed distribution can be explained by the hypothesized causal mechanism, then the distribution is said to be \textit{compatible} with said causal mechanism. In order to define compatibility rigorously, we first need to formally define the notion of a causal hypothesis.

    A hypothesis of causal mechanism is formally referred to as a \term{causal structure} and can be represented as a \term{directed acyclic graph}. A directed graph $\graph$ is an ordered tuple $\br{\nodes, \edges}$ of respectively \textit{nodes} and \textit{edges} where each edge $e \in \edges$ connects a \textit{pair} of nodes $n, m \in \nodes$ with a directed arrow $e = \bc{n \to m}$. A directed graph is \textit{acyclic} if there are no paths following the directions of the edges starting from and returning to the same node. The nodes $\nodes$ of a causal structure represent random variables while the edges $\edges$ represent a casual influence from one variable to another pursuant to the prescribed direction.

    Henceforth, we will utilize a number of familiar notions from graph theory and denote them accordingly. The \term{parents of a node} $n \in \nodes$ are all nodes which point directly into $n$, i.e. $\Pa[\graph]{n} \defined \bc{m \mid m \to n}$. Similarly defined are the \term{children of a node} $\Ch[\graph]{n} \defined \bc{m \mid n \to m}$. Recursively defined are the \term{ancestors of a node} $\An[\graph]{n} \defined \bigcup_{i\in\mathbb{N}} \Pa[\graph][i]{n}$ where $\Pa[\graph][i]{n} \defined \Pa[\graph]{\Pa[\graph][i-1]{n}}$ and $\Pa[\graph][0]{n} = n$ and the \term{descendants of a node} $\De[\graph]{n} \defined \bigcup_{i\in\mathbb{N}} \Ch[\graph][i]{n}$ where $\Ch[\graph][i]{n} \defined \Ch[\graph]{\Ch[\graph][i-1]{n}}$ and $\Ch[\graph][0]{n} = n$. Finally, we extend this notation to a subset of nodes $N \subseteq \nodes$ by performing a union over elements. As an example, the parents of the nodes $N \subseteq \nodes$ are denoted $\Pa[\graph]{N} = \bigcup_{n \in N} \Pa[\graph]{n}$.

    % A number of causal structures are depicted throughout this document (see \cref{fig:bell_scenario,fig:triangle_structure,fig:inflations}). As an example, the Triangle structure has $3$ observable nodes and $3$ latent nodes,
    % \[ \nodes_{O} = \bc{A, B, C} \qquad \nodes_{L} = \bc{X, Y, Z} \]
    % along with $6$ edges indicating that the latent nodes $\bc{X, Y, Z}$ cause and influence the observed nodes $\bc{A, B, C}$.
    % \[ \edges = \bc{\bc{X \to A}, \bc{Y \to A}, \bc{Y \to B}, \bc{Z \to B}, \bc{Z \to C}, \bc{X \to C}} \]

    Now that the notion of a causal structure $\graph$ is defined along with some standard graph-theoretic notation, it is now possible to precisely define the notion of compatibility between $\graph$ and a probability distribution $\prob[\nodes]$ defined over the nodes of $\graph$. Intuitively, the absence of certain causal connections within $\graph$ should limit the set of possible observations that can be made on $\graph$. Specifically, the causal structure $\graph$ hypothesizes that each variable $n \in \nodes$ is causally influenced by its parents $\Pa[\graph]{n}$, then $n$ should be independent of all other variables, except its own descendants when conditioned over $\Pa[\graph]{n}$. This is known as the \term{local Markov property}:\footnote{Here $A \indep B \mid C$ denotes statistical independence between $A$ and $B$ conditional on $C$ where $A,B,C \subseteq \nodes$, i.e. for a given probability distribution $\prob[\nodes]$, $\prob[AB \mid C] = \prob[A \mid C]\prob[B \mid C]$.}
    \[ n \indep \br{\nodes \setminus \De[\graph]{n}} \mid \Pa[\graph]{n} \quad \forall n \in \nodes \eq \label{eq:local_markov_property} \]
    If a given distribution $\prob[\nodes]$ defined over \textit{all} of the nodes $\nodes$ of $\graph$ satisfies \cref{eq:local_markov_property}, then $\prob[\nodes]$ is said to be \term{compatible with $\graph$}. Alternatively yet equivalently, the notion of compatibility can be interpreted as follows: a distribution $\prob[\nodes]$ is compatible with $\graph$ if $\prob[\nodes]$ can be consistently factorized into conditional probability distributions for each variable $n \in \nodes$ conditioned on its parents $\Pa[\graph]{n}$:
    \[ \prob[\nodes] = \prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] = \prob[n_1 \mid \Pa[\graph]{n_1}] \times \cdots \times \prob[n_k \mid \Pa[\graph]{n_k}] \eq \label{eq:local_markov_property_operation}\]
    The conditional distributions in \cref{eq:local_markov_property_operation} (i.e. $\{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes\}$) are referred to as a set of \term{causal parameters for $\graph$}. If a distribution $\prob[\nodes]$ violates \cref{eq:local_markov_property} or equivalently cannot be factorized according to \cref{eq:local_markov_property_operation}, $\prob[\nodes]$ is said to be \term{incompatible} with $\graph$. When one is specified with a joint distribution $\prob[\nodes]$ defined over all nodes of a causal structure $\graph$, it is trivially easy to verify whether or not $\prob[\nodes]$ is compatibility with $\graph$; \cref{sec:satisfaction_of_d_sep_relations} describes in detail how this is accomplished.

    A challenge is present when one is supplied with a partial observation, i.e. a joint distribution $\prob[\nodes_{O}]$ where $\nodes_{O} \subset \nodes$ is some subset of variables referred to as the \term{observable nodes} $\nodes_O$. In such cases, \cref{eq:local_markov_property_operation} can not be verified by direction calculation because it is certainly possible that the parents of some observable node are themselves unobservable $\Pa[\graph]{\nodes_{O}} \not \subseteq \nodes_{O}$. In such cases, compatibility between $\graph$ and $\prob[\nodes_{O}]$ depends on the existence or non-existence of a set of causal parameters for $\graph$ such that $\prob[\nodes_{O}] = \prod_{n \in \nodes_{O}} \prob[n \mid \Pa[\graph]{n}]$. The unobservable nodes are termed \term{latent nodes} $\nodes_L = \nodes \setminus \nodes_{O}$ representing hidden random variables that are hidden either by some fundamental process or cannot be measured due to other limitations.
    \todo[TC]{Explain why this is difficult more.}
    \todo[TC]{Figure out if I should include d-sep here}
    \todo[TC]{Explain what this work and inflation accomplishes regarding this}

    In practice however, checking causal parameters systemically is intractable, especially when the cardinality (number of outcomes) of the latent variables is unspecified. Instead, \term{causal compatibility inequalities}\footnote{We refer to these inequalities as ``causal compatibility inequalities'' instead of simply Bell inequalities for two reasons. Firstly, ``Bell inequalities'' usually are associated specifically with the Bell structure. Secondly, the inequalities derived in this work are fundamentally distinct from a typical Bell inequality in that these inequalities are \textit{polynomial} over $\prob[\nodes_{O}]$ instead of \textit{linear}.} are probabilistic inequalities defined over $\prob[\nodes_{O}]$ and its induced marginals which constraint the set of compatible probability distributions. If a probability distribution $\prob[\nodes_{O}]$ happens to \textit{violate} any causal compatibility inequality, then that distribution deemed \textit{incompatible}. In \cref{sec:deriving_inequalities}, a brief discussion of existing techniques for deriving these inequalities will take place.

    Unfortunately, a singular inequality cannot prove that a given distribution is \textit{compatible}, only \textit{incompatible}. Instead, a \textit{complete characterization} of compatibility would consist of a complete set of all valid causal compatibility inequalities; satisfaction of a complete set of inequalities does prove compatibility. Currently however, it is unknown how to obtain a complete characterization for any given causal structure, including the Triangle structure.

    From the perspective of identifying quantum non-classicality, a causal structure $\graph$ adopts the role of a classical hypothesis. Therefore, non-classicality becomes synonymous with incompatibility: if a distribution $\prob[\nodes_{O}]$ is incompatible, then it is non-classical. Henceforth, we will use these two terms interchangeably. From a resource standpoint, if the non-classicality of $\prob[\nodes_{O}]$ can be witnessed by an inequality $I$, but $\prob[\nodes_{O}]$ can be implemented using quantum states and measurements, then the causal compatibility inequality $I$ represents a physical task where quantum resources out-perform classical resources.

@@

@@

    \subsection{Satisfaction of d-Separation Relations}
    \label{sec:satisfaction_of_d_sep_relations}
    \todo[TC]{Absorb the relevant components of the next section into this one}
    \todo[TC]{Given a distribution and a causal structure, does the distribution satisfy the conditional d-Separation relations?}

    \subsection{Marginal Satisfiability and Inequalities}
    \label{sec:marginal_satisfiability}
    \label{sec:marginal_linear}\label{sec:linear_programs}

    Given a set of observable variables $\jointvar = \bc{v_1, \ldots, v_k}$ where each $v \in \jointvar$ is a statistical random variable, one can conceivably make a measurement over of these variables and obtain a joint probability distribution $\prob[\jointvar] = \prob[v_1,\ldots,v_n]$. More generally however, and depending on the context, it is possible to make a measurement on a set of the system $V \subset \jointvar$ in order to obtain a marginal distribution $\prob[V]$. With complete generality however, one could in principle be limited so as to only be able to record observations on portions of the system $\mscenario = \bc{V_1, \ldots, V_m \mid \forall i : V_i \subseteq \jointvar}$ obtaining a collection of marginal probability distributions $\bc{\prob[V_1], \ldots, \prob[V_m] \mid \forall i: V_i \in \mscenario}$.

    % For a causal structure $\graph$, a chosen specification of a conditional distributions for each variable  is referred to as a set of \term{causal parameters for $\graph$}:
    % \[  \eq \label{eq:causal_parameters} \]
    It is important to note that a set of causal parameters is not unique; the numerical nature of set of causal parameters is a \text{choice}. Altogether, a marginal model $\prob^{\mscenario}$ is \term{compatible} with a causal structure $\graph$ if there exists a choice of causal parameters such that each context $\prob[V] \in \prob^{\mscenario}$ can be \textit{recovered} from the following series of operations:
    \begin{enumerate}
        \item Obtain a joint probability distribution over \textit{all} variables of the causal structure using the chosen casual parameters:\footnote{Here $\prod$ denotes the product distribution $\prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] = $.}
        \item Marginalize over the latent variables of $\graph$ to obtain a joint distribution over the observed variables $\nodes_O$:
        \[ \prob[\nodes_O] = \sum_{\nodes_L} \prob[\nodes] \eq \label{eq:observable_marginalization} \]
        \item Finally marginalize over the observed nodes not in $V$ to obtain $\prob[V]$:
        \[ \prob[V] = \sum_{\nodes_O \setminus V} \prob[\nodes_O] \eq \label{eq:context_marginalization} \]
    \end{enumerate}
    It is important to note that compatibility between $\prob^{\mscenario}$ and $\graph$ requires that it is possible to recover \textit{each} context $\prob[V] \in \prob^{\mscenario}$ from the \textit{same} set of causal parameters. If it is not possible for $\prob^{\mscenario}$ to be recovered for \textit{any} choice of causal parameters, then $\prob^{\mscenario}$ is \term{incompatible}.
